---
output: 
  pdf_document: 
    keep_tex: yes
---


# Importing libraries : 
```{r}
library(tidyverse)
library(lubridate)
library(ggthemes)
library(YieldCurve)
```

# Data preparing :

## Importing Morocco rates data (daily) : 

```{r}
data <- read_delim("Moroccan rates.csv", 
    delim = ";", escape_double = FALSE, col_types = cols(Date = col_date(format = "%d/%m/%Y")), 
    trim_ws = TRUE)
```

```{r}
data_maroc = data %>% pivot_longer(cols=c(-Date),names_to = "x",values_to = "Y") %>% 
  mutate(x = as.numeric(x),Y = as.numeric(Y)/100)
```


## Lissing data :

In order to have same $x_k-x_{k-1}$ for all $k$ :  

```{r}
delta <- 3/12
newx <- seq(0, 30, by = delta)
newx[1] <- 	1/360

predictNS <- function(params, tau) {
  beta0 <- params[1]
  beta1 <- params[2]
  beta2 <- params[3]
  lambda <- params[4]
  
  term1 <- beta0
  term2 <- beta1 * (1 - exp(-lambda * tau)) / (lambda * tau)
  term3 <- beta2 * ((1 - exp(-lambda * tau)) / (lambda * tau) - exp(-lambda * tau))
  
  yield <- term1 + term2 + term3
  return(yield)
}

NS <- function(x, r) {
  params <- Nelson.Siegel(rate = r, maturity = x)
  rate <- predictNS(params, newx)
  return(rate)
}

mydata <- data_maroc %>%
  group_by(Date) %>%
  reframe(maturity = newx, Yliss = NS(x,Y))

write.csv(mydata, "dataliss.csv", row.names = FALSE)
```

## Import lissed data : 

```{r}
data <- read_csv("dataliss.csv")
colnames(data) = c("t","x","Y")
```


## Data we're using is Morocco weekly rates 2016~2018 after lissage using NS : 

```{r}
t0 = as.Date("2016-01-01")
rates = data %>% 
 # filter(t>=t0)  %>%
  mutate(t = as.numeric(difftime(t, t0, units = "days"))/365  ) %>%
  filter(Y>0.001)
```

```{r}
maturities = (rates %>% select(x) %>% distinct())$x
x0 = 1/360
rshort = (rates %>% filter(abs(x-x0)<x0/2) )$Y
```


## Adding P(t,T) : 

```{r}
rates = rates %>% 
  mutate( P  = exp(-x*Y) )
```


## Ading rt : 

```{r}
rates = rates %>% 
  mutate( r = rep(rshort, each = length(maturities) ) )
```

## Ading f(t,T) : 

```{r}
rates = rates %>% 
  mutate( f = ifelse(abs(x-x0)<x0,r,-(log(P) - lag(log(P)))/(x-lag(x)) ) ) 
```


# EDA and risk premuim : 

## Visualising $Y(t,T)$  : 

```{r}
ggplot(rates) +
  aes(x = t, y = Y, colour = x, group = x) +
  geom_line() +
  scale_color_gradient() +
  theme_bw()
```

```{r}
ggplot(rates) +
  aes(x = x, y = Y, colour = t, group = t) +
  geom_line() +
  scale_color_gradient() +
  theme_bw()
```

## Visualising Forward rates $f(t,T)$ : 

```{r}
ggplot(rates) +
  aes(x = t, y = f, colour = x, group = x) +
  geom_line() +
  scale_color_gradient() +
  theme_bw()
```

```{r}
ggplot(rates) +
  aes(x = x, y = f, colour = t, group = t) +
  geom_line() +
  scale_color_gradient() +
  theme_bw()
```

## Visualising Short rate $r_t$ :

```{r}
ggplot(rates) +
  aes(x = t, y = r) +
  geom_line(colour = "#112446") +
  theme_bw()
```
## Visualising ZCB prices $\ln P(t,T)$: 

```{r}
ggplot(rates) +
  aes(x = t, y = log(P), colour = x, group = x) +
  geom_line() +
  scale_color_gradient() +
  theme_bw()
```

## Risk Premuim :  

```{r}
dt = 1/360
BSparams = rates %>% select(t,x,P) %>% group_by(x) %>% 
  reframe( volatility = sqrt( (1/dt)* var( log(P/lag(P)) ,na.rm = TRUE)), return = mean(log(P/lag(P)),na.rm = TRUE)/dt + volatility^2/2 )
```

```{r}
sigma = sqrt(mean(BSparams$volatility^2))
mu = mean(BSparams$return)

qt = (mu - rates$r )/sigma
  
rates = rates %>% 
  mutate(q = qt)
```

```{r}
ggplot(rates) +
  aes(x = t, y = q) +
  geom_line(colour = "#112446") +
  theme_bw()
```


## Data we're gonna use : 
```{r}
rates
```


# Benchmark Models 

## Nilson Siegle 

Calibration : 
```{r}
NSParameters = Nelson.Siegel( rates$f, rates$x)

NSrates <- function(coef, t){
  beta0 = coef[,1]
  beta1 = coef[,2]
  beta2 = coef[,3]
  tau = coef[,4]
  return(beta0 + beta1 * (1 - exp(-t/tau))/(t/tau) + beta2 * ((1 - exp(-t/tau))/(t/tau) - exp(-t/tau))) 
}
```

Calcul des estimations : 
```{r}
curveNS  = tibble( maturity = maturities, rates = NSrates(NSParameters,maturities))
```

Visualisation de courbe : 
```{r}
ggplot(curveNS) +
  aes(x = maturity, y = rates) +
  geom_line(colour = "#112446") +
  theme_bw()
```


## Svensson  

Calibration : 
```{r}
SParameters = Svensson( rates$f, rates$x)

Srates <- function(coef, t){
  beta0 = coef[,1]
  beta1 = coef[,2]
  beta2 = coef[,3]
  beta3 = coef[,4]
  tau1 = coef[,5]
  tau2 = coef[,6]
  term1 <- beta0
  term2 <- beta1 * (1 - exp(-t/tau1)) / (t/tau1)
  term3 <- beta2 * ((1 - exp(-t/tau1))/(t/tau1) - exp(-t/tau1))
  term4 <- beta3 * ((1 - exp(-t/tau2))/(t/tau2) - exp(-t/tau2))
  
  return(term1 + term2 + term3 + term4)
}
```

Calcul des estimations : 
```{r}
curveS  = tibble( maturity = maturities, rates = Srates(SParameters,maturities))
```

Visualisation de courbe : 
```{r}
ggplot(curveS) +
  aes(x = maturity, y = rates) +
  geom_line(colour = "#112446") +
  theme_bw()
```


## Comparaison des models benchmark : 

```{r}
ggplot(rates) +
  aes(x = x, y = f, colour = t, group = t) +
  geom_line() +
  geom_line(  aes(x = x, y =  NSrates(NSParameters,x) ),linewidth = 1.2, colour = "orange" )+
  geom_point(  aes(x = x, y =  NSrates(NSParameters,x) ),size = 1.6, colour = "orange" )+
  geom_line(  aes(x = x, y =  Srates(SParameters,x) ),linewidth = 1.2, colour = "green" )+
  geom_point(  aes(x = x, y =  Srates(SParameters,x) ),size = 1.6, colour = "green" )+
  theme_bw()
```

# Proposed Model (Three Factors Model TF) :

```{r}
# Defining initial Values :
delta = 0.25
N = length(maturities)
time = rates$t
f0 = (rates %>% filter(t==0) )$f

dt  = 1/360
timesDaily  = unique(rates$t)
dtW = 1/52
timeWeekly  = numeric(length(timesDaily)*dt/dtW)
for(i in 0:length(timeWeekly)){
  timeWeekly[i+1] = timesDaily[1+i*floor(dtW/dt)]
}


rates = rates %>% filter( t %in% timeWeekly )
```

```{r}
# Function for predictions : 
Fonc <- function(t,dt,r,q,params){
  N = length(r)
  lambda0 = params[1]
  lambda1 = params[2]
  lambda2 = params[3]
  kappa0  = params[4]
  kappa1  = params[5]
  kappa2  = params[6]
  
  sigma0  <- function(t, s, x) {
    return(lambda0 * exp(-kappa0 * (t  +x- s)))
  }
  mu0     <- function(t, s, x) {
    term1 <- exp(-kappa0 * (t  +x-s))
    term2 <- exp(kappa0 * s) - exp(-kappa0 * (t  +x- 2 * s))
    return(lambda0^2 * term1 * (term2 / kappa0))
  }
  sigma1  <- function(t, s, x) {
    return(lambda1 * exp(-kappa1 * (t + x-s)))
  }
  mu1     <- function(t, s, x) {
    term1 <- exp(-kappa1 * (t  +x- s))
    term2 <- exp(kappa1 * s) - exp(-kappa1 * (t +x- 2 * s))
    return(lambda1^2 * term1 * (term2 / kappa1))
  }
  sigma2  <- function(t, s, x) {
    return(lambda2 * exp(-kappa2 * (t +x- s)))
  }
  mu2     <- function(t, s, x) {
    term1 <- exp(-kappa2 * (t  +x- s))
    term2 <- exp(kappa2 * s) - exp(-kappa2 * (t  +x- 2 * s))
    return(lambda2^2 * term1 * (term2 / kappa2))
  }
  f         <- function(t, l, c, r, dW) {
    term1 <- c + ((r - l) / (2 * delta) + mu0(t,t,c) * c^2 + mu1(t,t,c) * ((r - l) / (2 * delta))^2 + mu2(t,t,c) * ((r - 2 * c + l) / delta^2)^2) * dt
    term2 <- sigma0(t,t,c) * c * dW[1] + sigma1(t,t,c) * ((r - l) / (2 * delta)) * dW[2] + sigma2(t,t,c) * ((r - 2 * c + l) / delta^2) * dW[3]
    result <- term1 + term2
    return(result)
  }
  

  rnext = r 
  dW = c(0,0,0) + rep(q*dt,3)/3
  for( k in 2:(N-1) ){
    rnext[k] = f(t, r[k-1], r[k], r[k+1], dW)
  }
  rnext[1] = f(t, r[1] , r[1], r[1+1], dW)
  rnext[N] = f(t, r[N-1] , r[N], r[N] , dW)
  
  return(rnext)
}


FoncAlea <- function(t,dt,r,q,params){
  N = length(r)
  lambda0 = params[1]
  lambda1 = params[2]
  lambda2 = params[3]
  kappa0  = params[4]
  kappa1  = params[5]
  kappa2  = params[6]
  
  sigma0  <- function(t, s, x) {
    return(lambda0 * exp(-kappa0 * (t  +x- s)))
  }
  mu0     <- function(t, s, x) {
    term1 <- exp(-kappa0 * (t  +x-s))
    term2 <- exp(kappa0 * s) - exp(-kappa0 * (t  +x- 2 * s))
    return(lambda0^2 * term1 * (term2 / kappa0))
  }
  sigma1  <- function(t, s, x) {
    return(lambda1 * exp(-kappa1 * (t + x-s)))
  }
  mu1     <- function(t, s, x) {
    term1 <- exp(-kappa1 * (t  +x- s))
    term2 <- exp(kappa1 * s) - exp(-kappa1 * (t +x- 2 * s))
    return(lambda1^2 * term1 * (term2 / kappa1))
  }
  sigma2  <- function(t, s, x) {
    return(lambda2 * exp(-kappa2 * (t +x- s)))
  }
  mu2     <- function(t, s, x) {
    term1 <- exp(-kappa2 * (t  +x- s))
    term2 <- exp(kappa2 * s) - exp(-kappa2 * (t  +x- 2 * s))
    return(lambda2^2 * term1 * (term2 / kappa2))
  }
  f         <- function(t, l, c, r, dW) {
    term1 <- c + ((r - l) / (2 * delta) + mu0(t,t,c) * c^2 + mu1(t,t,c) * ((r - l) / (2 * delta))^2 + mu2(t,t,c) * ((r - 2 * c + l) / delta^2)^2) * dt
    term2 <- sigma0(t,t,c) * c * dW[1] + sigma1(t,t,c) * ((r - l) / (2 * delta)) * dW[2] + sigma2(t,t,c) * ((r - 2 * c + l) / delta^2) * dW[3]
    result <- term1 + term2
    return(result)
  }
  

  rnext = r 
  dW = rnorm(3,mean = 0,sd = sqrt(dt)) + rep(q*dt,3)/3
  for( k in 2:(N-1) ){
    rnext[k] = f(t, r[k-1], r[k], r[k+1], dW)
  }
  rnext[1] = f(t, r[1] , r[1], r[1+1], dW)
  rnext[N] = f(t, r[N-1] , r[N], r[N] , dW)
  
  return(rnext)
}
```



```{r}
time = unique(rates$t) 
maxt = max(time)


# Defining the norm of vector : 
norm2 <- function(x){
  sum(x^2)
}

suivant <- function(tt,time){
  inext = 1
  while( time[inext]<tt ){
    inext=inext+1
  }
  return(time[inext+1])
}


# Define the mean squared error function
mse <- function(params) {
  mse_value = 0
  tt = time[1]
  i = 0 
  while(tt<maxt){
    r =  as.vector(rates %>% filter(t==tt) %>% select(f))$f
    q =  as.vector(rates %>% filter(t==tt) %>% select(q))$q[1]
    dt = suivant(tt,time)-tt
    prediction <- Fonc(tt,dt,r,q,params)
    tt = tt+dt
    r =  as.vector(rates %>% filter(t==tt) %>% select(f))$f
    mse_value <- mse_value + norm2( (r-prediction)/r ) 
  }
  return(mse_value/length(time))
}
```


```{r}
## Initial params :
params = c(
    lambda0 = 0.04157374,
    lambda1 = 0.01830442,
    lambda2 = 0.0926567,
    kappa0  = 0.09287434,
    kappa1  = 0.01661409,
    kappa2  = 0.0436558
          )

paramsoptim = c(
    lambda0 = 0.008742619 ,
    lambda1 = 12.13609 ,
    lambda2 = 0.02129127 ,
    kappa0  = 5.564819 ,
    kappa1  = 16.173 ,
    kappa2  = 6.821594
)
```


```{r}
# Record timing :
start_time <- Sys.time()

# Define the optimization algorithm
resultat <- optim(par = paramsoptim , fn = mse)

params = resultat$par

paste0("mse = ", round(sqrt(resultat$value)*100,2),"%")

cat(
  " Parametres estimed : \n
    lambda0 =",paramsoptim[1], ",
    lambda1 =",paramsoptim[2], ",
    lambda2 =",paramsoptim[3], ",
    kappa0  =",paramsoptim[4], ",
    kappa1  =",paramsoptim[5], ",
    kappa2  =",paramsoptim[6]
)

# Record end time
end_time <- Sys.time()

# Calculate elapsed time
elapsed_time <- end_time - start_time

# Print the result
cat("\n Time for running algorithm:", as.numeric(elapsed_time), "seconds\n")
```

## Backtesting for $h = 1$ 

```{r}
tt = time[23]
tt = time[1]
tt
r = as.vector(rates %>% filter(t==tt) %>% select(f))$f
N = length(r)
q =  as.vector(rates %>% filter(t==tt) %>% select(q))$q[1]
ttsuivante =  suivant(tt,time)
dt= ttsuivante-tt

freal =  as.vector(rates %>% filter(t==ttsuivante) %>% select(f))$f

freallag =  as.vector(rates %>% filter(t==tt) %>% select(f))$f

tt = tt+dt
festim =  Fonc(tt,dt,r = r,q=q, params = paramsoptim)
```

```{r}
CurveQ = tibble(
  x = maturities ,
  freal = freal, 
  freallag = freallag,
  festim = festim,
  
  festimNS = curveNS$rates,
  festimS = curveS$rates
  )
```


```{r}
CurveQ %>% 
  pivot_longer(cols = c(freal,festim,freallag,festimNS,festimS),names_to = "Method", values_to = "f(t,T)") %>% 
  ggplot() +
    aes(x = x, y = `f(t,T)`, colour = Method, group = Method) +
    geom_line() +
    scale_color_hue(direction = 1) +
    theme_bw()
```

```{r}
## Erreurs d'estimations : 
cat(
  "\n",
  "Erreur NS  =", 1000*norm2(freal - CurveQ$festimNS), "%\n",
  "Erreur Sv  =", 1000*norm2(freal - CurveQ$festimS) , "%\n",
  "Erreur lag =", 1000*norm2(freal - CurveQ$freallag) , "%\n",
  "Erreur TF  =", 1000*norm2(freal - CurveQ$festim)  , "%\n"
)
```


## Backtesting for $h>1$ : 

```{r}
tt = time[1]
tt
h = 8
r = as.vector(rates %>% filter(t==tt) %>% select(f))$f
N = length(r)
q =  as.vector(rates %>% filter(t==tt) %>% select(q))$q[1]

freallag =  as.vector(rates %>% filter(t==tt) %>% select(f))$f

for(i in 1:h){ 
  dt = suivant(tt,time) - tt 
  q =  as.vector(rates %>% filter(t==tt) %>% select(q))$q[1]
  festim =  Fonc(tt,dt,r = r,q=q, params = paramsoptim)
  tt = tt+dt
  r = festim
}

freal =  as.vector(rates %>% filter(t==tt) %>% select(f))$f
```

```{r}
CurveQ = tibble(
  x = maturities ,
  freal = freal, 
  freallag = freallag,
  
  festim = festim,
  
  festimNS = curveNS$rates,
  festimS = curveS$rates
  )
```


```{r}
CurveQ %>% 
  pivot_longer(cols = c(freal,festim,freallag,festimNS,festimS),names_to = "Method", values_to = "f(t,T)") %>% 
  ggplot() +
    aes(x = x, y = `f(t,T)`, colour = Method, group = Method) +
    geom_line() +
    scale_color_hue(direction = 1) +
    theme_bw()
```


```{r}
## Erreurs d'estimations : 
cat(
  "\n",
  "Erreur NS  =", 1000*norm2(freal - CurveQ$festimNS), "%\n",
  "Erreur Sv  =", 1000*norm2(freal - CurveQ$festimS) , "%\n",
  "Erreur lag =", 1000*norm2(freal-  CurveQ$freallag) , "%\n",
  "Erreur TF  =", 1000*norm2(freal - CurveQ$festim)  , "%\n"
)
``` 




## Simulation of Forward rates : 

```{r}
Sim = tibble(
  t = c(), x = c(), f = c()
)

tt = 0 
for(i in 1:30){
  r = as.vector(rates %>% filter(t==tt) %>% select(f))$f
  N = length(r)
  q =  as.vector(rates %>% filter(t==tt) %>% select(q))$q[1]
  ttsuivante =  suivant(tt,time)
  dt= ttsuivante-tt
  festim =  FoncAlea(tt,dt,r = r,q=q, params = paramsoptim)
  
  Sim = Sim %>% bind_rows(
    tibble(
    t = rep(tt,N), x = maturities, f = festim
        )
  )
  tt = tt+dt
}
```



```{r}
ggplot(Sim) +
  aes(x = t, y = f, colour = x, group = x) +
  geom_line() +
  scale_color_gradient() +
  theme_bw()
```

